```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

---
title: "Practical Machine Learning"
output: html_document
---
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.

In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

The goal of this project is to predict the manner in which the participants did the exercise. This is the “classe” variable in the training set.

Data Loading
------------
```{r}
library(caret)

#Set seed for repeating process
set.seed(44422)
#Downloading file 1
fileUrl<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
fileLocal<-"pml-training.csv"
download.file(fileUrl,destfile=fileLocal,method="curl")
trainData <- read.csv(fileLocal,na.strings=c("NA",""));
#Downloading file2
fileUrl<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
fileLocal<-"pml-testing.csv"
download.file(fileUrl,destfile=fileLocal,method="curl")
testData <- read.csv(fileLocal,na.strings=c("NA",""));

```
Exploratory analysis and data cleaning
------------------------------------------------
```{r}
#Now deleting columns with excesive NAs
trainReductionData<-trainData[, which(colSums(is.na(trainData))==0)]

#Erasing first columns. No relevant for modeling
trainReductionData<-trainReductionData[-c(1:7)]
```

Creating partitions
--------------------

```{r}
#Partition of data
inTrain <- createDataPartition(y=trainReductionData$classe,p=0.75,list=FALSE)
traindatafinal = trainReductionData[inTrain,]
validationdatafinal =trainReductionData[-inTrain,]
```
Now we explore the model for selecting the best predictors
```{r results='hide'}
#Selecting variables for model 
#We limit number of trees for computational time issues
#We test with full number of trees and output
modelFit <- train(classe ~ ., data = traindatafinal, method="rf",importance=TRUE,do.trace=1,ntree=20)
```

```{r}
#important variables
library(plyr)
importVar <- varImp(modelFit,type=2)
importVar<-importVar$importance
importVar$vars<-rownames(importVar)
newVar<-arrange(importVar,desc(importVar$Overall))
```
We select top 15 important variables
```{r}
newVar[1:15,]
newVar<-newVar[c(1:15),2]
#adding classe for model fit
newVar[length(newVar)+1]<-"classe"

#We take the first 15 top important variables 
trainDataImportant<-traindatafinal[,newVar]

```

New model whith new predictors
-------------------------------
```{r results='hide'}

#Training model with new data
modelFit <- train(classe ~ ., data = trainDataImportant, method="rf",importance=TRUE,do.trace=1,ntree=50)
```
cross-validation and estimation of error
----------------------------------------
```{r}
#Data validation
validationDataImportant<-validationdatafinal[,newVar]
finalValidation<- confusionMatrix(validationDataImportant$classe, predict(modelFit, validationDataImportant))
finalValidation
```

Finally the last test with the original test data set.
```{r}
#Data for testing prediction
finalTest <- testData[,newVar[1:15]]
submit<-predict(modelFit, finalTest)
submit
```


The data for assigment two is generated
```{r}
answers <- as.vector(submit)

pml_write_files = function(x) {
   n = length(x)
   for (i in 1:n) {
      filename = paste0("problem_id_", i, ".txt")
      write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, 
                  col.names = FALSE)
   }
}

pml_write_files(answers)

```
